\section{Introduction}

In this paper, we discuss a partial differential equation that models the evolution in time of an age-structured population. Our equation is useful for age-structured models, demographic models where the population at time $t$ has an age distribution superimposed on it. In other words, at a given time $t$, the ages of the individuals in a population are also considered. Age-structured models are related to general physiologically-structured models, where any other variable (such as size or weight) can replace the age variable. For example, the following \href{https://www.jstor.org/stable/3547254?seq=1}{paper} uses a physiologically structured population model to model a ``size-structured consumer population feeding on a non-structured prey population", studying the qualitative and quantitative population dynamics of a planktivorous fish population. Consequently, the ideas we discuss in this paper, though focused on age, can also be extended to other physiological structures. \\


Anderson Gray McKendrick (A.G. McKendrick) was a Scottish physician and epidemiologist. In 1926, McKendrick published a paper titled \textit{Applications of mathematics to medical problems}. He aimed to study the transfer of disease caused by interactions between people and apply mathematical modeling to epidemiology; his paper introduced a version of the McKendrick-von Forester equation. \cite{mck1925} McKendrick's paper went relatively unseen; independently, in 1959, biophysics Heinz von Foerster discovered the same equation when studying cell divisions. \cite{keyfitz_keyfitz} The equation itself was later named in recognition of their independent work, as what we now know as the \textbf{McKendrick–von Forester equation}. \\

The \textbf{McKendrick–von Forester equation} is a form of the advection equation that incorporates the mortality rate at some age to study the population density at a given age and time. \\

In our project, we will derive the general age-structured model in Logan's \textit{Applied Partial Differential Equations} \cite{logan}:
\[ u_t = -u_a-m(a)u, \, \, \, a>0, \, t>0\]
\[ u(0,t) = \int_0^{\infty}b(a,t)u(a, t) \, da, \, \, \, t>0\]
\[ u(a, 0)=f(a), \, \, \, a \geq 0\]

where the newly introduced terms $b(a, t)$ and $f(a)$ represent the average reproduction rate and the population of female organisms, respectively. \\

We will then discuss the stable age structure to see what happens over a long time. Finally, we plan on presenting the renewal equation (a simplified version of the age-structured model above where $b = b(a)$ and $m =$ constant), which we will solve using the method of characteristics. \\

In conclusion, we will discuss various extensions of the McKendrick-von Forester equation and its importance in helping scientists understand population models. 
